---
title: "Homework 2_Data Wrangling_jm4998"
author: "Junhui Mi"
date: "2019/9/24"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

_Read data from Excel_

```{r read_excel, warning = FALSE}
Mr_data = read_excel("../dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = 1, range = "A2:N336") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls))
```

_Join 2017 and 2018 precipitation data_

```{r join_tables, warning = FALSE}
prec_2018 = read_excel("../dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = 3, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2018, 12), month = month.name) %>%
  drop_na(total) 

prec_2017 = read_excel("../dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = 4, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2017, 12), month = month.name) %>%
  drop_na(total)

prec_all = full_join(prec_2018, prec_2017)
```

## Conclusions

(1) Mr_data shows Mr. Trash Wheel data, with `r length(pull(Mr_data, dumpster))` observations. There are `r length(Mr_data)` key variables: `r colnames(Mr_data)`.

(2) prec_all shows the combination of 2017 precipitation and 2018 precipitation data, with `r length(pull(prec_all, month))`. observations. There are `r length(prec_all)` key variables: `r colnames(prec_all)`.

(3) The the total precipitation in 2018 is `r sum(pull(filter(prec_all, year == 2018), total))`.

(4) The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(Mr_data, year == 2017), sports_balls))`.

