---
title: "Homework 2_Data Wrangling_jm4998"
author: "Junhui Mi"
date: "2019/9/28"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

_Read data from Excel and clean Mr_data_

```{r read_excel, warning = FALSE}
Mr_data = read_excel("../dataset/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = 1, range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls))
```

_Join 2017 and 2018 precipitation data_

```{r join_tables, warning = FALSE}
prec_2018 = read_excel("../dataset/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2018, 12), month = month.name[as.integer(month)]) %>%
  drop_na(total) 

prec_2017 = read_excel("../dataset/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2017, 12), month = month.name[as.integer(month)]) %>%
  drop_na(total)

prec_all = full_join(prec_2018, prec_2017)
```

## Description

(1) "Mr_data" shows Mr. Trash Wheel data, with `r length(pull(Mr_data, dumpster))` observations. There are `r length(Mr_data)` variables: `r colnames(Mr_data)`; "dumpster" and "sports_balls" are key variables.

(2) "prec_all" shows the combination of 2017 precipitation and 2018 precipitation data, with `r length(pull(prec_all, month))`. observations. There are `r length(prec_all)` key variables: `r colnames(prec_all)`.

(3) The the total precipitation in 2018 is `r sum(pull(filter(prec_all, year == 2018), total))`.

(4) The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(Mr_data, year == 2017), sports_balls))`.

# Problem 2

_Read and clean pols_month, unemployment and snp data from csv_

```{r clean_csv, warning = FALSE}
pols_month = read_csv(file = "../dataset/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.name[as.integer(month)]) %>%
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    values_to = "value") %>% 
  filter(value == 1 | value == 2) %>% 
  select(-day, -value) %>%
  mutate(
    president = replace(president, president == "prez_gop", "republican"),
    president = replace(president, president == "prez_dem", "democrat"))
  

unemployment = read_csv(file = "../dataset/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  rename(January = jan, February = feb, March = mar, April = apr,
         May = may, June = jun, July = jul, August = aug,
         September = sep, October = oct, November = nov, December = dec) %>%
  pivot_longer(
    cols = January:December,
    names_to = "month",
    values_to = "unemployment_rate") %>%
  select(year, month, everything())

snp = read_csv(file = "../dataset/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year"), sep = "/") %>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(year, month, close)
```

_Left join three datasets by date_

```{r merge_data}
pols_month_2 = pols_month %>%
  mutate(date = paste(year, month, sep = "-")) %>%
  select(date, gov_gop:president)

unemployment_2 = unemployment %>%
  mutate(date = paste(year, month, sep = "-")) %>%
  select(date, unemployment_rate)

snp_2 = snp %>%
  mutate(date = paste(year, month, sep = "-")) %>%
  select(date, close)

pols_snp = left_join(pols_month_2, snp_2, by = "date")
pols_snp_unemployment = left_join(pols_snp, unemployment_2, by = "date") %>%
  separate(date, c("year", "month"), sep = "-")
```

## Description

(1) The three initial datasets:  "pols_month" contains 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time; "unemployment" contains 68 observations of 13 variables related to unemployment rate; "snp" contains 787 observations of 2 variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole.

(2) The resulting dataset "pols_snp_unemployment" merges data from "pols_month", "unemployment" and “snp”, with `r length(pull(pols_snp_unemployment, month))` rows and `r length(pols_snp_unemployment)` columns. It contains data from year `r min(pull(pols_snp_unemployment, year))` to `r max(pull(pols_snp_unemployment, year))`.Its variables are: `r colnames(pols_snp_unemployment)`. "year" and "month" are key variables which merge into "date" to be more accurate.

# Problem 3

_Read and clean baby names from csv_

```{r name_csv, warning = FALSE}
baby_names = read_csv(file = "../dataset/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(childs_first_name = str_to_title(childs_first_name),
         ethnicity = recode(ethnicity, 
                            "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "BLACK NON HISP" = "BLACK NON HISPANIC",
                            "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  distinct() 
```

The "baby_names" dataset collects `r length(pull(baby_names, year_of_birth))` popular baby name samples from year `r min(pull(baby_names, year_of_birth))` to `r max(pull(baby_names, year_of_birth))`.

_Popularity of the name “Olivia”_

```{r olivia, warning = FALSE}
baby_names_olivia = baby_names %>%
  filter(childs_first_name == "Olivia", gender == "FEMALE") %>%
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>%
  knitr::kable(format = "html")
baby_names_olivia
```

The tibble "baby_names_olivia" above shows the rank in popularity of the name “Olivia” as a female baby name over time.

_The most popular male name_

```{r male_names, warning = FALSE}
popular_male_names = baby_names %>%
  filter(gender == "MALE", rank == "1") %>%
  select(year_of_birth, ethnicity, childs_first_name) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>%
  knitr::kable(format = "html")
popular_male_names
```

The tibble "popular_male_names" above shows the most popular name among male children over time.

_Scatter plot of the popularity of names for male, white non-hispanic children born in 2016_

```{r create_plot, warning = FALSE}
plot_data = baby_names %>%
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", 
         year_of_birth == "2016")
ggplot(plot_data, aes(x = rank, y = count)) + 
  geom_point() +
  ggtitle("Scatter plot of male, white non-hispanic children born in 2016")
ggsave("Scatter plot of male, white non-hispanic children born in 2016.png")
```

The scatter plot above shows the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016.